### 2023-02-15 Code/initial_implementation.Rmd
Initial Implementation

```{r}
set.seed(1234)
mystat <- function(x) {
    c(c(x),  # Bootstrap will return actual pseudo-sample, mean, and standard deviation
      mean(x),
      sd(x))
}
```

```{r}
n <- 100
phi <- 0.5
blksize <- 6
B = 100

mysim <- function(n, phi, blksize, B) {
  if (phi == 0) {
    x <- rnorm(n)
  } else {
    x <- arima.sim(list(ar = phi), n = n) * sqrt(1 - phi^2)
  }
  bts <- boot::tsboot(x, mystat, l = blksize, sim = "fixed", R = B)
  list_cdf <- matrix(, nrow = 0, ncol = n)
  list_theta <- matrix(, nrow = 0, ncol = 2)
  sum_theta <- 0
  for (i in 1:B) {
    list_cdf <- rbind(list_cdf, c(ecdf(bts$t[i, (1:n)])(sort(bts$t0[1:n]))))
    list_theta <- rbind(list_theta, c(bts$t[i, -(1:n)]))
  }
  exp_cdf <- colMeans(list_cdf)
  exp_theta <- colMeans(list_theta)
  ks_values <- c()
  for (i in 1:B) {
    ks_values <- c(ks_values, max(abs(sqrt(n) * (list_cdf[i,] - exp_cdf - pnorm(sort(bts$t0[1:n]), list_theta[i, 1], list_theta[i, 2]) + pnorm(sort(bts$t0[1:n]), exp_theta[1], exp_theta[2])))))
  }
  c(ks_values)
}

```

```{r}
print(mysim(n, phi, blksize, B))
```