### 2023-09-16 code/simulation.Rmd
Renamed Code

```{r}
set.seed(1234)
mystat <- function(x, dist) {
    c(c(x),  # Bootstrap will return actual pseudo-sample, mean, and standard deviation
      MASS::fitdistr(x, dist)$estimate)
}
```

```{r}
mysim <- function(n, blksize, B, dist, f, phi = 0, theta = 0) {
  if (mean(phi == 0)) {
      x <- arima.sim(list(ma = theta), n = n) / sqrt(1 + sum(theta^2))
  } else {
      x <- arima.sim(list(ar = phi, ma = theta), n = n) * sqrt(1 - sum(phi^2) / 1 + sum(theta^2))
  }
  emp <- ecdf(x)(sort(x)) # observed empirical distribution
  fit_theta <- MASS::fitdistr(x, dist)$estimate # observed fitted values
  fit <- do.call(f, c(list(sort(x)), as.list(fit_theta))) # observed fitted distribution
  obsv_ks <- sqrt(n) * max(abs(emp - fit)) # observed ks statistic
  bts <- boot::tsboot(x, mystat, l = blksize, sim = "fixed", R = B, dist = dist)
  list_cdf <- matrix(, nrow = 0, ncol = n) 
  # initialize list of each pseudo-sample's ecdf applied to original sample values
  list_theta <- matrix(, nrow = 0, ncol = 2) 
  # initialize list of each pseudo-sample's fitted theta
  # bts$t[i, (1:n)] : values of bootstrap sample
  # bts$t[i, -(1:n)] : theta (mu, and sigma)
  sum_theta <- 0
  for (i in 1:B) {
    list_cdf <- rbind(list_cdf, c(ecdf(bts$t[i, (1:n)])(sort(x))))
    list_theta <- rbind(list_theta, c(bts$t[i, -(1:n)]))
  }
  exp_emp_cdf <- colMeans(list_cdf)
  exp_theta <- colMeans(list_theta)
  exp_fit_cdf <- do.call(f, c(list(sort(x)), as.list(exp_theta)))
  bias_term <- exp_emp_cdf - exp_fit_cdf
  ks_values <- c()
  for (i in 1:B) {
    emp_cdf <- list_cdf[i,]
    fit_cdf <- do.call(f, c(list(sort(x)), as.list(list_theta[i, ])))
    ks_values <- c(ks_values, sqrt(n) * max(abs((emp_cdf - fit_cdf - bias_term))))
  }
  c(mean(ks_values > obsv_ks))
}

```

```{r ar = -0.3}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- -0.3
theta <- 0
sim_ar1_n.3 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_n.3, "../Data/sim_ar1_n.3.csv", row.names = FALSE)
hist(sim_ar1_n.3)
```

```{r ar = -0.2}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- -0.2
theta <- 0
sim_ar1_n.2 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_n.2, "../Data/sim_ar1_n.2.csv", row.names = FALSE)
hist(sim_ar1_n.2)
```

```{r ar = -0.1}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- -0.1
theta <- 0
sim_ar1_n.1 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_n.1, "../Data/sim_ar1_n.1.csv", row.names = FALSE)
hist(sim_ar1_n.1)
```

```{r ar = 0}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- 0
theta <- 0
sim_ar1_0 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_0, "../Data/sim_ar1_0.csv", row.names = FALSE)
hist(sim_ar1_0)
```

```{r ar = .1}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- .1
theta <- 0
sim_ar1_.1 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_.1, "../Data/sim_ar1_.1.csv", row.names = FALSE)
hist(sim_ar1_.1)
```

```{r ar = .2}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- .2
theta <- 0
sim_ar1_.2 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_.2, "../Data/sim_ar1_.2.csv", row.names = FALSE)
hist(sim_ar1_.2)
```

```{r ar = .3}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- .3
theta <- 0
sim_ar1_.3 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_.3, "../Data/sim_ar1_.3.csv", row.names = FALSE)
hist(sim_ar1_.3)
```

```{r ar = .5}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- .5
theta <- 0
sim_ar1_.5 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar1_.5, "../Data/sim_ar1_.5.csv", row.names = FALSE)
hist(sim_ar1_.5)
```

```{r ma = .5}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- 0
theta <- .5
sim_ma1_.5 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ma1_.5, "../Data/sim_ma1_.5.csv", row.names = FALSE)
hist(sim_ma1_.5)
```

```{r ar = .5, ma = .3}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- .5
theta <- .3
sim_arma <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_arma, "../Data/sim_arma.csv", row.names = FALSE)
hist(sim_arma)
```

```{r ar = (.5, .3)}
nrep = 1000
n = 100
blksize = 5
B = 1000
dist = 'normal'
f = pnorm
phi <- c(.5, .3)
theta <- 0
sim_ar2 <- replicate(nrep, mysim(n, blksize, B, dist, f, phi, theta))
write.csv(sim_ar2, "../Data/sim_ar2.csv", row.names = FALSE)
hist(sim_ar2)
```