### 2024-09-02 manuscript/statistical_practice_cover_letter.tex

\documentclass{article}
\usepackage{charter}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr,graphicx}
\usepackage{xspace}

\fancypagestyle{firstpage}{%
  \fancyhf{} % Clear header/footer
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}
\fancypagestyle{otherpages}{%
  \fancyhf{}% Clear header/footer
  \renewcommand{\headrulewidth}{1pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}

\AtBeginDocument{\thispagestyle{firstpage}}
\pagestyle{otherpages}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

\begin{document}

\vspace*{\dimexpr-\headsep-\headheight-1pt}

% \includegraphics[width=1in]{logo.jpg}

\rule{\linewidth}{1pt}

\bigskip

%----------------------------------------------------------------------------------------
%	YOU ONLY NEED TO FILL IN THE FOLLOWING LATEX MACROS
%----------------------------------------------------------------------------------------

% Title of the paper
\newcommand{\PaperTitle}{Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test\xspace}

% Full name of the authors
\newcommand{\AuthorOne}{Mathew Chandy\xspace}
\newcommand{\AuthorTwo}{Elizabeth D. Schifano\xspace}
\newcommand{\AuthorThree}{Jun Yan\xspace}
\newcommand{\AuthorFour}{Xianyang Zhang\xspace}

% Full name of the Journal's Editor in Chief
\newcommand{\EditorInChief}{Min-ge Xie\xspace}

% Journal
\newcommand{\Journal}{The American Statistician\xspace}

% Body
\newcommand{\Contribution}{X\xspace}
\newcommand{\Problem}{Y\xspace}
\newcommand{\Solution}{Z\xspace}
\newcommand{\Evaluation}{D\xspace}
\newcommand{\NoveltyClaim}{E\xspace}

% List of contributions
\newcommand{\ContributionOne}{Contribution A.\xspace}
\newcommand{\ContributionTwo}{Contribution B.\xspace}
\newcommand{\ContributionThree}{Contribution C.\xspace}
\newcommand{\ContributionFour}{Contribution D.\xspace}

% Suggested editors
\newcommand{\EditorOne}{Editor \#1\xspace}
\newcommand{\EditorOneExpertise}{A\xspace}

\newcommand{\EditorTwo}{Editor \#2\xspace}
\newcommand{\EditorTwoExpertise}{B\xspace}

\newcommand{\EditorThree}{Editor \#3\xspace}
\newcommand{\EditorThreeExpertise}{C\xspace}

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\hfill
\begin{tabular}{ l @{} }
  \today \\[12pt] % Date
  Department of Statistics,\\
  University of Connecticut\\
  Room 323, Philip E. Austin Building\\
  215 Glenbrook Road, Unit 4120\\
  Storrs, CT 06269-4120\\
  \\
  Department of Statistics and Data Science,\\
  University of California, Los Angeles\\
  8125 Math Sciences Bldg. Box 951554\\
  Los Angeles, CA 90095-1554\\
  \\
  Department of Statistics,\\
  Texas A\&M University\\
  447 Blocker Building (BLOC)\\
  3143 TAMU | 155 Ireland St\\
  College Station, TX 77843-3143
\end{tabular}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\bigskip

Dear Dr. Xie,

\bigskip

We are pleased to submit the manuscript entitled ``Nonparametric Block Bootstrap 
Kolmogorov-Smirnov Goodness-of-Fit Test" for 
consideration in the Statistical Practice section of 
\textit{American Statistician}. We certify that this 
manuscript has neither been published nor is under consideration by another 
journal.

Our paper introduces a novel approach
which can widen the applications of the Kolmogorov--Smirnov (KS) test, a 
goodness-of-fit test which relies on the data
being serially dependent and the parameters for the hypothesized distribution
being specified.

This paper introduces a bias correction for the test based on the non-parametric 
block bootstrap that accounts for serial dependence and does not require
specification of parameters.
The main contribution is a new methodological development which is useful
for addressing a broad array of
practical problems in financial analysis, environmental science, as 
well as other situations that require goodness-of-fit tests on serially
dependent data. An application of our method to test the distribution
of S\&P 500 index stock returns is shown in the manuscript.

This article will prove to be a valuable expansion to statistical practice.
Please feel free to reach out to us if you need any other information concerning 
this manuscript.

\bigskip

Sincerely yours,

\vspace{50pt}

Mathew Chandy\\ \\On behalf of Elizabeth D. Schifano, Jun Yan, and 
Xianyang Zhang

\end{document}