### 2024-09-17 manuscript/cov.tex
cover letter for General section of TAS

\documentclass[12pt]{article}
\usepackage{charter}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr,graphicx}
\usepackage{xspace}

\fancypagestyle{firstpage}{%
  \fancyhf{} % Clear header/footer
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}
\fancypagestyle{otherpages}{%
  \fancyhf{}% Clear header/footer
  \renewcommand{\headrulewidth}{1pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}

\AtBeginDocument{\thispagestyle{firstpage}}
\pagestyle{otherpages}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

\begin{document}

\vspace*{\dimexpr-\headsep-\headheight-1pt}

% \includegraphics[width=1in]{logo.jpg}

\rule{\linewidth}{1pt}

\bigskip

%----------------------------------------------------------------------------------------
%	YOU ONLY NEED TO FILL IN THE FOLLOWING LATEX MACROS
%----------------------------------------------------------------------------------------

% Title of the paper
\newcommand{\PaperTitle}{Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test\xspace}

% Full name of the authors
\newcommand{\AuthorOne}{Mathew Chandy\xspace}
\newcommand{\AuthorTwo}{Elizabeth D. Schifano\xspace}
\newcommand{\AuthorThree}{Jun Yan\xspace}
\newcommand{\AuthorFour}{Xianyang Zhang\xspace}

% Full name of the Journal's Editor in Chief
\newcommand{\EditorInChief}{Min-ge Xie\xspace}

% Journal
\newcommand{\Journal}{The American Statistician\xspace}

% Body
\newcommand{\Contribution}{X\xspace}
\newcommand{\Problem}{Y\xspace}
\newcommand{\Solution}{Z\xspace}
\newcommand{\Evaluation}{D\xspace}
\newcommand{\NoveltyClaim}{E\xspace}

% List of contributions
\newcommand{\ContributionOne}{Contribution A.\xspace}
\newcommand{\ContributionTwo}{Contribution B.\xspace}
\newcommand{\ContributionThree}{Contribution C.\xspace}
\newcommand{\ContributionFour}{Contribution D.\xspace}

% Suggested editors
\newcommand{\EditorOne}{Editor \#1\xspace}
\newcommand{\EditorOneExpertise}{A\xspace}

\newcommand{\EditorTwo}{Editor \#2\xspace}
\newcommand{\EditorTwoExpertise}{B\xspace}

\newcommand{\EditorThree}{Editor \#3\xspace}
\newcommand{\EditorThreeExpertise}{C\xspace}

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\hfill
\begin{tabular}{ l @{} }
  \today \\[12pt] % Date
  Department of Statistics and Data Science,\\
  University of California, Los Angeles\\
  8125 Math Sciences Bldg. Box 951554\\
  Los Angeles, CA 90095-1554
\end{tabular}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\bigskip

Dear Dr. Xie,

We are pleased to submit our manuscript entitled 
``Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit
Test"  for consideration in the \emph{General} Section of the
\textit{American Statistician}. We certify that this manuscript has
neither been published nor is under consideration by any other
journal.

Our paper addresses a gap in the application of the Kolmogorov-Smirnov
(KS) test for assessing the marginal goodness-of-fit of hypothesized
distributions in the context of serially dependent stationary
observations. While parametric block bootstrap procedures typically
require specification of the serial dependence structure, a
nonparametric block bootstrap procedure is advantageous as it avoids
this requirement. The nonparametric block bootstrap version of the KS
test extends the nonparametric approach studied by Basu and Rao
(2004) to cases of serial dependence. However, such a version does not
yet exist, and our paper fills this gap. The proposed test has wide
applications in time series modeling, particularly in scenarios where
marginal distributions are of interest, such as in financial time series.

We believe that the methodological developments in this manuscript are
of broad interest and are generally not highly technical, making it a
suitable fit for the \emph{General} Section of your journal.

Thank you for your consideration, and we look forward to your feedback.

\bigskip

Sincerely yours,

\vspace{2cm}

Mathew Chandy\\ \\On behalf of Elizabeth D. Schifano, Jun Yan, and 
Xianyang Zhang

\end{document}