### 2025-10-08 manuscript/supplement.tex
figures fixed

%&latex
\documentclass[12pt]{article}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL
\usepackage{comment}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{setspace}

\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\mc}[1]{\textcolor{green}{(MC: #1)}}
\newcommand{\xz}[1]{\textcolor{cyan}{(XZ: #1)}}

\allowdisplaybreaks

\usepackage[pagewise]{lineno}
\linenumbers*[1]
% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
        \expandafter\let\csname old#1\expandafter\endcsname\csname 
        #1\endcsname
        \expandafter\let\csname oldend#1\expandafter\endcsname\csname 
        end#1\endcsname
        \renewenvironment{#1}%
        {\linenomath\csname old#1\endcsname}%
        {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
        \patchAmsMathEnvironmentForLineno{#1}%
        \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
        \patchBothAmsMathEnvironmentsForLineno{equation}%
        \patchBothAmsMathEnvironmentsForLineno{align}%
        \patchBothAmsMathEnvironmentsForLineno{flalign}%
        \patchBothAmsMathEnvironmentsForLineno{alignat}%
        \patchBothAmsMathEnvironmentsForLineno{gather}%
        \patchBothAmsMathEnvironmentsForLineno{multline}%
}

\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\bf Supplementary Material to\\
  ``Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test''
}
\if0\blind
{
  \author[1,2]{Mathew Chandy}
  \author[1]{Elizabeth D. Schifano}
  \author[1]{Jun Yan\thanks{
    The authors gratefully acknowledge the National Science Foundation (NSF).}\hspace{.2cm}}
  \author[3]{Xianyang Zhang$^{\ast}$ }
  \affil[1]{Department of Statistics, University of Connecticut}
  \affil[2]{Department of Statistics and Data Science, University of California, Los Angeles}
  \affil[3]{Department of Statistics, Texas A\&M University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \author{Anonymous Author}
} \fi

\maketitle

\section{Comparison of Two Bias Corrections}

A heuristic justification for the $K_n(x)$ bias correction term was
provided in the main manuscript. It may also help to demonstrate
empirically that this the block bootstrap KS test with the $K_n$ bias
correction holds its size at least as well as the block boostrap KS
test with the \citet{babu2004goodness} bias correction and is at
least as powerful. For this reason, we conducted a simulation study
with parameter settings identical to the main study in the paper
except for the choice of bias correction.

\begin{figure}[tbp]
  \includegraphics[width = .9\textwidth]{figures/normal_C_n}
  \centering
  \vspace{-10pt}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal normal distribution with true data generating distribution
    being $N(8,8)$ when using \citet{babu2004goodness}'s bias correction term.}
  \label{fig:qq_n_C_n}
    \hspace{3cm}
  \includegraphics[width = .9\textwidth]{figures/gamma_C_n}
  \vspace{-5pt}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal gamma distribution with true data generating distribution
    being $\Gamma(8,1)$ when using \citet{babu2004goodness}'s bias correction term.}
  \label{fig:qq_g_C_n}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/rr_C_n}
  \caption{Empirical power curve as a function of $\tau$ for
    $n \in \{100, 200, 400, 800\}$ and true marginal distribution
    $\in \{N(8,8), \Gamma(8,1)\}$, when using \citet{babu2004goodness}'s bias correction term. When the data was generated from $N(8,8)$,
    we tested for the gamma family. When the data was generated from
    $\Gamma(8,1)$, we tested for the normal family.
  }
  \label{fig:rr_C_n}
\end{figure}

Based on Figures~\ref{fig:qq_n_C_n, fig:qq_g_C_n, fig:rr_C_n}, 
the empirical performance in both size and power analysis is very similar under 
either choice of bias correction 
term. This suggests that the circular block bootstrap already corrects for
much of the temporal dependence, and that the choice of centering
has a very small effect on residual bias. Using $C_n(x)$ has the advantage
of requiring slightly less computation, but we 
still recommend $K_n(x)$ for its theoretical qualities, but acknowledge that
performance may be indistinguishable in practical application, especially
with large sample size and dependence that decays rapidly.

\section{Block Size Selection}

The choice of block size is a key factor influencing the performance
of block bootstrap methods, particularly in settings with serial
dependence. While our main simulations used a fixed block size
$l = \lceil n^{1/3} \rceil$, the optimal block size depends on the
stength and structure of dependence \citep{hall1995blocking,
  buhlmann1999block,  politis2004automatic}. Motivated by this, we
further examined the use of an automatic block-length selection scheme
as proposed by \citet{politis2004automatic}, which aims to adaptively
choose the block size based on the data. The simulation settings
remain the same as in the study reported earlier.

% To address this, we considered  using the automatic block-length selection
% scheme for circular block bootstrap as described in Section~3.3 of
% \citet{politis2004automatic}. Unfortunately, using this setting results in
% the selection of block
% sizes that can be quite large (sometimes as large or larger than $n$), leading
% to less variation between the bootstrapped
% samples. Figures~\ref{fig:large_block}
% demonstrate that for highly temporally
% dependent samples (especially when the autocorrelation is negative),
% the block size is frequently larger than $\sqrt{n}$. For more moderate 
% positive temporal dependence, the block size may not be overly large
% as frequently. Still, even for these moderately dependent samples, the 
% block sizes are almost always larger than $l = \lceil n^{1/3} \rceil$, a setting
% which was considered optimal by \citep{buhlmann1999block}. This
% is demonstrated in Figure~\ref{fig:block_dist} with a histogram
% of the selected block sizes for the 20000 samples with $n = 800$, $\tau = 0.5$, 
% and $X_i \sim N(8, 8)$.

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/block_dist}
  \caption{Distribution of the selected block sizes using
  \citet{politis2004automatic}'s procedure applied to 20000 length-800 time 
  series with
    Kendall's $\tau = 0.5$ data generating distribution $N(8,8)$. The vertical
    dashed line is $l = \lceil n^{1/3} \rceil$ for $n = 800$, which is equivalent 
    to 10.
  }
  \label{fig:block_dist}
% \end{figure}

% \begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/large_block}
  \caption{Proportion of selected block sizes using the method of
  \citet{politis2004automatic} that are larger than
    $\sqrt{n}$ for $n \in \{100, 200, 400, 800\}$,
    $\tau \in \{-0.75, -0.50, -0.25, 0, 0.25, 0.50, 0.75\}$, and true
    marginal distribution $N(8,8)$ and $\Gamma(8,1)$.
  }
  \label{fig:large_block}
\end{figure}

Automatic block size selection tends to yield larger block sizes than our
fixed choice $l = \lceil n^{1/3} \rceil$, as expected, but under strong
negative or very strong positive dependence it systematically produces
block sizes that are excessively large, sometimes as large as or larger
than $n$. Note that in our simulations, we set the final block size to 
$\min \{n, \hat l\}$, where
$\hat l$ is the block size selected by \citet{politis2004automatic}'s
procedure. Figure~\ref{fig:block_dist} illustrates the typical behavior by
showing the distribution of block sizes selected by \citet{politis2004automatic}
for $n = 800$, $\tau = 0.5$, and $X_i \sim N(8,8)$. The majority of selected
block sizes exceed $\lceil n^{1/3} \rceil$, and many approach or exceed
$\sqrt{n}$. Figure~\ref{fig:large_block} presents the proportion of selected
block sizes exceeding $\sqrt{n}$ across varying levels of dependence and sample 
sizes.
The selection of block sizes larger than $\sqrt{n}$ occurs systematically under 
strong
negative dependence, and is also frequent under very strong positive dependence.

\begin{figure}[tbp]
  \includegraphics[width = .9\textwidth]{figures/alt_normal}
  \centering
  \vspace{-10pt}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal normal distribution with true data generating distribution
    being $N(8,8)$, when using \citet{politis2004automatic}'s procedure
    to select block sizes.}
  \label{fig:alt_qq_n}
    \hspace{3cm}
  \includegraphics[width = .9\textwidth]{figures/alt_gamma}
  \vspace{-5pt}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal gamma distribution with true data generating distribution
    being $\Gamma(8,1)$, when using \citet{politis2004automatic}'s procedure
    to select block sizes.}
  \label{fig:alt_qq_g}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/alt_rr}
  \caption{Empirical power curve as a function of $\tau$ for
    $n \in \{100, 200, 400, 800\}$ and true marginal distribution
    $\in \{N(8,8), \Gamma(8,1)\}$, when using \citet{politis2004automatic}'s 
    procedure
    to select block sizes. When the data was generated from $N(8,8)$,
    we tested for the gamma family. When the data was generated from
    $\Gamma(8,1)$, we tested for the normal family.
  }
  \label{fig:alt_rr}
\end{figure}

We evaluated the size and power performance of the NPBB test under automatic
block length selection. The size performance is shown
in Figures~\ref{fig:alt_qq_n} and \ref{fig:alt_qq_g}, which display Q-Q plots
of the $p$-values under the null for Normal and Gamma margins, respectively.
Recall from the Q-Q plots in the main paper that the cubic root 
setting holds its size quite well with small $n$ for $|\tau| \leq 0.25$,
whereas achieving nominal size for $|\tau| \geq 0.5$ require
$n > 200$, and for $|\tau| \geq 0.75$ require $n > 800$. The automatic
selection shows similar limitations for $|\tau| > 0.5$, where it does
not hold its size. Moreover, deviations from the diagonal line appear
larger for negative $\tau$, likely due to the high proportion of
occurrence of overly large block sizes. For example, while the cubic
root setting held its size well at $\tau = -0.25$ with small~$n$, the
automatic selection does not. At the same time, under $\tau = 0$ or
$\tau = 0.25$, the method with block size selected by
\citep{politis2004automatic} maintains reasonable size for moderately
positively dependent series.

Figure~\ref{fig:alt_rr} shows the power results under automatic selection.
When compared to the power curve in the main paper, the two settings exhibit 
similar power
for $|\tau| \leq 0.5$, with the automatic selection sometimes yielding higher
rejection rates, particularly for smaller negatively dependent samples.
Overall, while the automatic selection from \citep{politis2004automatic}
performs reasonably well in certain settings, we recommend caution when using
it under negative~$\tau$ and smaller sample sizes, where maintaining nominal
size may be more difficult. In such cases, it may be sensible to apply a more
restrictive rule when setting the final block size---for example, using
$\min \{\sqrt{n}, \hat{l}\}$.

\bibliographystyle{asa}
\bibliography{citations}
\end{document}